---
title: "STAT 413/613: HW on List Columns and  COVID19"
author: "Your Name"
date: "`r Sys.Date()`"
output:
  html_document:
    number_sections: yes
    theme: cerulean
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align  = "center",
                      fig.height = 5, 
                      fig.width  = 6)
```

# Instructions {-}
1. Clone this homework repo to your homework directory as a new repo.
2. Rename the starter file under the analysis directory as `hw_01_yourname.Rmd` and use it for your solutions.   
3. Modify the "author" field in the YAML header.  
4. Stage and Commit R Markdown and HTML files (no PDF files).   
5. **Push both .Rmd and HTML files to GitHub**.   
- Make sure you have knitted to HTML prior to staging, committing, and pushing your final submission.  
6. **Commit each time you answer a part of question, e.g. 1.1**   
7. **Push to GitHub after each major question**   
8. When complete, submit a response in Canvas  
    
- Only include necessary code to answer the questions.
- Most of the functions you use should be from the tidyverse. Unnecessary Base R or other packages not covered in class will result in point deductions.
- Use Pull requests and or email to ask me any questions. If you email, please ensure your most recent code is pushed to GitHub.

- **Learning Outcome**
  + Use tidyverse functions to create, clean, tidy, and manipulate data frames in a list column
  + Apply purrr functions when working with list columns
  + Employ joins to manipulate data from multiple data frames

- **Context** 
  + This assignment looks at COVID-19 data based on the most recent data as of the date you do the work.

# Scoring Rubric {-}



# Load global and US confirmed cases and deaths data into a nested data frame
1. Create a variable called `url_in` to store this URL: "https://github.com/CSSEGISandData/COVID-19/tree/master/csse_covid_19_data/csse_covid_19_time_series"
```{r}
library(tidyverse)
library(tibble)
library(dplyr)
library(httr) 
library(broom)
library(purrr)

url_in<-("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/")

```

2. Create a tibble named `df` with a variable called `file_names` with a row for each of the following four file names to be loaded from the URL:
    + time_series_covid19_confirmed_global.cs
    + time_series_covid19_deaths_global.csv
    + time_series_covid19_confirmed_US.csv
    + time_series_covid19_deaths_US.csv
```{r}
df<-tibble(
  file_names=c("time_series_covid19_confirmed_global.csv","time_series_covid19_deaths_global.csv","time_series_covid19_confirmed_US.csv","time_series_covid19_deaths_US.csv")
)
```
    
3. Create a variable in the data frame called `url` that puts `url_in` on the front of each file_name to create a complete URL.
```{r}
df %>% 
  mutate(url=str_c(url_in,file_names,sep = ""))->df


```

4. Use `mutate()` with `map()` to create a list column called `data` with each row holding the downloaded data frame for each file name
```{r}
df %>% 
  mutate(data=map(url,~read_csv(.)))->df1
```


5. Add a factor variable to `df` called `"`case_types`"` with the **unique** portions of the file names.
```{r}
df1 %>% 
  mutate(case_ty=str_replace(df1$file_names,"time_series_covid19_",""),
         case_types=str_replace(case_ty,".csv","")) %>% 
  select(-case_ty)->df2
head(df2)
```

6. Remove any columns other than `case_types` and `data` from `df`.
- `df` should have four observations of two variables.
```{r}
df2$file_names<-NULL
df2$url<-NULL
```

 
# Clean Data  
1. Use `map()` to add the names from each of the four data frames to a new variable in `df` called `vars` and visually compare them to identify issues.
```{r}
df2$vars<-map(df2$data,~names(.))


```
 
 

2. Take the following steps to fix any issues and create consistent data frames.  
a. Create a short helper function called `fix_names()` which takes three arguments, a data frame, a string, and a replacement pattern. It should replace all occurrences of the string in the names of the variables in the data frame with the replacement pattern.
```{r}

fix_names<-function(data,str,rep){
 names(data[[1]])[names(data[[1]])==str]=rep
 names(data[[2]])[names(data[[2]])==str]=rep
 names(data[[3]])[names(data[[3]])==str]=rep
 names(data[[4]])[names(data[[4]])==str]=rep

 data
  
}

```

a. Convert "Province/State" and "Country/Region" to "Province_State" "Country_Region" .
```{r}
df2$data<-fix_names(df2$data,"Province/State","Province_State")
df2$data<-fix_names(df2$data,"Country/Region","Country_Region")

```

b. Convert "admin2 to "County" and "Long_" to "Long".
```{r}
df2$data<-fix_names(df2$data,"Admin2","County")
df2$data<-fix_names(df2$data,"Long_","Long")

```

c. Remove the variables "UID", "iso2", "iso3", "code3", "FIPS", and "Combined_Key" from the US data.
```{r}
rm34<-c("UID", "iso2", "iso3", "code3", "FIPS","Combined_Key")

map_if(df2$data,~all(rm34%in%colnames(.)), ~select(.,-"UID", -"iso2", -"iso3", -"code3", -"FIPS",-"Combined_Key"))->df2$data
```

d. Add variables `Population` and `County` to the data frames where missing.
```{r}
map_if(df2$data,~!"Population"%in%colnames(.), ~mutate(.,Population=NA))->df2$data
map_if(df2$data,~!"County"%in%colnames(.), ~mutate(.,County=NA))->df2$data

```

e. Add a variable called `Country_State` that combines the country with the province/state while keeping the original columns.
```{r}
map(df2$data, ~mutate(.,Country_State=str_c(.$Country_Region,.$Province_State,sep = "_")))->df2$data


```

f. Update the values in `df$vars` when complete to check for consistency.
- Hint: Look at help for `map_if()`
```{r}
df2$vars<-map(df2$data,~names(.))

```


# Tidy each dataframe 
1. Use `map()` along with pivot_longer to tidy each data frame and as part of the pivot, ensure the daily values are in a variable called "Date" and use a lubridate function inside the pivot to ensure it is of class date.
2. Save the new data frame to a variable called `df_long`
```{r}
library(lubridate)
df3<-df2
df3 %>% 
  mutate(data=map(df3$data,~pivot_longer(.,cols=-c("Province_State", "Country_Region", "Lat", "Long","County","Population", "Country_State"),names_to = "Date",values_to="cases",names_transform=list(Date=mdy))))->df_long
```




# Add Continents 
1.  Use `map()` to add a new variable called `Continent` to each data frame.  
- Hint: use the package {countrycode} to get the continents.
- If you don't have it already, use the console to install. 
- Then load package countrycode and look at help for `countrycode::countrycode`
- You will get some warning messages about NAs which you will fix next.

```{r}
library(countrycode)
 map(df_long$data,~mutate(.,Continent=countrycode(Country_Region,origin = 'country.name', destination = 'continent')))->df_long$data
 

```

# Fix NAs for Continents
- Use `map()` with `case_when()` to replace the NAs due to "Diamond Princess", "Kosovo", "MS Zaandam" with the most appropriate continent
```{r}

  map(df_long$data,~mutate(.,Continents=case_when(
                              Country_Region=="Diamond Princess"     ~"Europe",
                              Country_Region=="Kosovo"     ~"Europe",
                              Country_Region=="MS Zaandam"     ~"Europe",
                              TRUE          ~as.character(Continent))) )->df_long$data
map(df_long$data,~select(.,-Continent))->df_long$data

```

- Use `map()` with `unique()` to confirm five continents in the global data frames and one in the US data frames
```{r}
map(df_long$data,~unique(.$Continents))

```
 
# Unnest the Data Frames    
1. Unnest and ungroup the data frame `df_long` and save into a new data frame called `df_all
```{r}
df_long %>% 
  unnest(data) %>% 
  ungroup()->df_all
```

2. Remove original `df` and `df_long` dataframes from the environment
```{r}
rm(df)
rm(df_long)
```

3. Remove the `vars` variable from df_all
```{r}

df_all %>% 
  select(-vars)->df_all
```



# Get World Population Data
1.  Read in World population data for 2019 into its own data frame called `df_pop`
-   Use the provided CSV or you can go to the [UN source](https://population.un.org/wpp/Download/Standard/CSV/)
  + The CSV has a few changes in country names to match the COVID data, e.g., US, and Iran.
  + Note: the UN data is in thousands so it can have fractional values
```{r}
read.csv("../hw_07-yiyao-wang/data/WPP2019_TotalPopulation.csv")->df_pop
```
  
2. Use a join to remove all Locations that are not in the `df_all` data frame.
```{r}
names(df_pop)[2]="Country_Region"
semi_join(df_pop,df_all,by="Country_Region")->df_pop1
```

3. Add the ranks for each location for population and population density to `df_pop`
```{r}
df_pop1 %>% 
  mutate(rank_p=rank(-PopTotal),
        rank_d=rank(-PopDensity) )->df_pop2
```

# Add Population Data to `df_all`
- Use a join to add the data from `df_pop` to `df_all`
- This means there will be two columns with population data:
  + `Population` for US Counties
  + `TotPop` for the country level
```{r}
full_join(df_all,df_pop2,by="Country_Region") ->df_all1
```  

# Analyse Data
1. Create a data frame by with data grouped by `Country_Region`, `Continent` `case_type`, `rank_p` and `rank_d` that summarizes the current totals and the totals as a percentage of total population.
  - Be sure to look at how the data is reported so the numbers make sense.
```{r}
df_all1 %>% 
  group_by(Country_Region, Continents,case_types, rank_p , rank_d) %>% 
 summarize(curr_t=max(cases),
            curr_p=((curr_t/1000)/PopTotal)*100)->curr_all 

```
  
2. What are the 20 Countries with the most confirmed cases and what is the percentage of their total population affected?
```{r}
 
curr_all[!duplicated(curr_all, fromLast=TRUE), ] %>% 
  group_by(case_types) %>% 
mutate(rank_confirm=round(rank(-curr_t))) ->rank_all
   filter(rank_all,case_types=="confirmed_global") %>% 
arrange(rank_confirm) %>% 
  head(20)->t20_c
t20_c

```

3. What are the 20 Countries with the most deaths and what is the percentage of their total population affected?
```{r}
rank_all %>% 
   filter(case_types=="deaths_global") %>% 
arrange(rank_confirm) %>% 
  head(20)->t20_d
t20_d
```

4. Try to interpret the results by just looking at the rankings for the totals with the rankings for total population and population density.

the result shows that the higher the rank for total population, the more confirmed cases and deaths. This is due to the large population base resulting in an increase in the absolute number of infections.

# Which countries in the top 20 for percentage of population affected are Not in the top 20 for the absolute number of cases and deaths?
- Try to interpret the results by just looking at the rankings for the totals with the rankings for total population and population density.
```{r}
rank_all %>%
 filter(case_types=="confirmed_global") %>% 
  arrange(-curr_p) %>% 
  head(20)->affect_c
  
rank_all %>%
 filter(case_types=="deaths_global") %>% 
  arrange(-curr_p) %>% 
  head(20)->affect_d

anti_join(affect_c,t20_c,by="Country_Region")
anti_join(affect_c,t20_d,by="Country_Region")

```

Although these countries are among the highest in terms of proportion of the population affected, they generally have a small population, which results in a small absolute number of cases of diagnosis or death.

# Create two plots, one for the number of cases and one for the number of deaths over time for the top 20 country/states faceting by continent. 
- Use appropriate scales for the axes.
- Create two sets of plots
- Interpret each plot.
```{r}
library(ggplot2)
t20_c[[1]]->t20_confname
t20_d[[1]]->t20_deatname
df_all %>% 
filter(Country_Region %in%t20_confname,case_types=="confirmed_global") %>%
  filter(is.na(Province_State)==TRUE) %>%
  ggplot(aes(x=Date,y=cases,color=Country_Region))+
  geom_line()+
  facet_wrap(~Continents,scales = "free")

# the plot shows with the exception of the United States, most of the top 20 countries with confirmed cases increased significantly around June. The number of confirmed cases in the United States has risen sharply since March.
```


```{r}
df_all %>% 
filter(Country_Region%in%t20_deatname,case_types=="deaths_global") %>% 
  filter(is.na(Province_State)==TRUE) %>% 
ggplot(aes(x=Date,y=cases,color=Country_Region))+
  geom_line()+
  facet_wrap(~Continents,scales = "free")

# The images of the number of deaths and confirmed cases are basically the same. But in most European countries there was a slow plateau between June and October. The death toll then rose sharply. This indicates a second outbreak in winter.
```
 

